\chapter{Literature review}

\section{Introduction}
This chapter will provide an overview of different topics that informed design
decisions made during the initial stages of the project. We will start looking
at the discipline of Open Source Intelligence, which refers to all those
protocols and techniques used by government and private agencies to piece
together intelligence reports using publicly available sources, since Linkero is
a tool that facilitates the structured collection and analysis of a limited
portion of open source data. The legal aspect of data collection and analysis
will be explored with a reasoned summary of what is GDPR and how it impacts
similar online services. Three sections will be dedicated to industry level best
practices in relation to the security of servers facing the public internet, the
usability of software interfaces and current guidelines to build reusable,
mantainable and expandable code. Finally we will review briefly what are some of
the current services already offering brand monitoring tools and how they differ
from one another.


\section{Open Source Intelligence}
Open Source Intelligence (more commonly referred to as OSINT) is a relatively
young discipline, that is concerned with the art of piecing together strategic
intelligence from public sources of information. Michael Bazzel, a leading OSINT expert,
defines it as:
\begin{displayquote}
any intelligence produced from publicly available information that is collected,
exploited, and disseminated in a timely manner to an appropriate audience for
the purpose of addressing a specific intelligence requirement. [\ldots] For the
CIA, it may mean information obtained from foreign news broadcasts. For an
attorney, it may mean data obtained from official government documents that are available to the
public. For most people, it is publicly available content obtained from the
internet \cite{MB15}.
\end{displayquote}

As Bezzell explains, OSINT is not necessarialy based on online sources, at least
in its most broad definition. Journalistic style, old fashion dossiers filled
with newspaper clippings are a form of OSINT. However it is fair to say that
whenever OSINT is mentioned today, it will automatically produce the expectation
that a large proportion of content is sourced through the internet. Another
important author in this field, Stewart K. Bertram, explains:
\begin{displayquote}
older OSINT research was limited by both the coverage of its information and the
ability of the researcher to focus the capability on a specific subject, be it
a person, location or topic. [\ldots] What has changed this status quo is the
arrival of the Internet, and particularly the explosion in the use of social
media technology circa 2000. The rise of these two technologies created a
multilingual, geographycally distributed, completely unregulated publishing
platform to which any user could also become an author and a publisher. [\ldots]
By increasing the coverage and focus of OSINT the Internet effectively promoted
OSINT from a supporting role to finally sit alongside other more clandestine and
less accessible investigative capabilities \cite{SB15}.
\end{displayquote}

This explosion of sources of information causes another problem: reliability.
Unless we are sourcing information from a scientifical magazine which
follows the rigorous fact checking protocol of most scientific researches, then
we are facing a vast landscape of information with various degrees of
veridicity: reliable fact-checked sources on one side and \emph{fake news} at
the other end of the spectrum. The work of the OSINT investigator is to move in
this virtually infinite universe of news, pick only relevant information and
establish how reliable they are. Again this is not new, it is called
\emph{intelligence analysis}: ``[\ldots] the application of individual and
collective cognitive methods to weigh data and test hypoteses within a secret
socio-cultural context'' \cite{JH07}. Information
are scored from A (reliable) down to E (unreliable), plus F (reliability
unknown).

Within the domain of counterfeit investigations, OSINT can have a number of
roles to play. First off and foremost, should be used to establish if the items
being sold are genuine or counterfeits. Secondly, it can be used to establish
the extent of profits made (and therefore loss of income for the original brand)
by the merchant selling them. And finally, OSINT can provide vital help in
identifying the identity of the merchant as well as that of the manifacturer.


\section{GDPR}
The General Data Protection Regulation (EU) 2016/679 (a.k.a. GDPR) is the latest
legislative effort of the Europea Union, that regulates the collection
and use of European citizens' data. The two main goals that guided this new
legislation were already set in 2015 by the Council of the European Union, which
stated:
\begin{displayquote}
The twofold aim of the Regulation is to enhance data
protection rights of individuals and to improve business opportunities by facilitating the free
flow of personal data in the digital single market \cite{CoEU956515}.
\end{displayquote}
Essentially the GDPR gives more control to European citizens over the use of
their data, thus restricting the ability of private and public organizations to
process Personally Identifiable Information (a.k.a. PII) data, and at the same
time it becomes a legislative framework that standardizes personal data
processing in every country member of the European Union. Let's see in more
detail what this entails and what are the consequences for a product like
Linkero.

First off it is important to clarify some of the terminology used in the
legislation. \emph{Personally identifiable information}, or PII, is defined in
privacy laws as one single piece of data that alone can identify a single person. For
instance: the first and last name of a person, an email address, a physical
address, a phone number, passport serial number, credit card number, a picture portraying an
individual, a combiantion of username and password. GDPR applys to the
collection, processing and storage of PII. A \emph{Data Controller} is any
organization that collects and manages PII belonging to European citizens. That
includes online private companies (e.g. Google, Facebook the most obvious), as well as
public institutions like primary and secondary schools and hospitals to name
few. Another party that is involved in the handling of private data is the
\emph{Data Processor}, which in some cases conicides with the the Data
Controller itself, when data is analysed internally, or with any other organization (e.g.
contractors, vendors, suppiers, etc.) that receives PII from the Data Controller
and processes it in its behalf. Finally the \emph{Data Protection Officer} is a
public office that is appointed to overlook and audit every Data Controller in
its area of control.

What GDPR means for European citizens is that as of May 25\textsuperscript{th}
2018, when it officially came into force, it grants much more rights over their
own personal data, compared to previous legislations. These rights include
\emph{data portability}, which is to say that any Data Controller has to
structure the collection and storage of PII in such a way that those information
can be removed, transferred and re-allocated to another Data Controller without
any further modification. In other words, this is a way to guarantee
interoperability between Data Controllers, preventing single Controllers to
lock-in their users. GDPR also states that EU citizens have the \emph{right to
information and transparency}, so that at any moment they can request their PII
to be disclosed as they have been collected up to that point in time. Another
right that generated a lot of discussion, is the \emph{right to be forgotten}, which
states that users can have their information permanently erased from a Data
Controller, if they no longer require their services.

GDPR also sets more stringent constraints to the operation of organizations
that fall within its scope. GDPR is self-declared \emph{transnational in scope},
meaning that it applys to every company, anywhere in the world, as long as they hold EU
citizens' PII data. Obviously, enforcement is an issue in cases such as
companies that are based somewhere in the Cayman Islands. However this aspect is
mainly ment at large multinational corporations, that could simply
transfer their databases to another jurisdiction outside the EU in an attempt
to circumvent its reach. Under the current definition of GDPR, those
multinationals would still be subjected to EU law and would still be
accountable as long as they have legal presence in any EU member state,
regardless of where the PII data actually sits. GDPR also establishes a stronger
\emph{Data Subject consent}. This means that Data Controllers can no longer
assume that their users' data can be processed and used for whatever purpose,
just because PII were given to them during the registration process. Nor they
can pre-tick the data consent acknoledgment box, or ask the  user to tick a box
only if they do not consent to the treatment of their data. On the contrary,
Data Controllers have to explicitly ask and receive users' consent, and consent
has to be renewed over time. Finally, GDPR requires organization to \emph{report
data breaches} within 72 hours after the breach was discovered. Data Controllers
not only have to notify the supervisory authority and their users, but the
notification has to carry adequate details to explain  how the incident
occurred, how many users are affected, and how is the organization going to respond.

With all these new rights and obligations in place, the last part of the
legislation concernes to the \emph{sanctions} for organizations that fail to
comply. There are three levels of penalties that can be imposed:
\begin{enumerate}
  \item a warning letter for the first time an organization is found in breach
  of GDPR, and if the breach is non-intentional;
  \item a more serious breach could result in the order to commit to regular
  periodic data protection audits;
  \item and finally, if the breach is deemed to be serious enough, GDPR the
  prescribes two sets of financial fines depending of which obligations were
  missed by the Data Controller, the harsher of which could amount to up to
  \euro20 millions or up to 4\% of the annual turnover estimated from the
  prior financial year.
\end{enumerate}

The question now is: how does GDPR affect online services that for their
nature collect and process bits of PII that are publicly available? First and
foremost, GDPR applys only to PII belonging to EU citizens, therefore we can say
that an online service that collects PII can keep on providing its functions
without any further thought as long as it stays clear of EU citizens But let's
say some users will be pointing the service to PII of European users, will that
be sufficient to be found in breach of GDPR? The answer to this latter question
is not a clear-cut \emph{no}, there are actually use-cases where collection can
still be an option. Obviously those cases do not include when Data Subjects
who expressed consent to their PII to be collected and stored by anyone online,
that is not how data scraping for intelligence purpose works, especially in the
context of open source investigations. Although it is too early for similar
cases to have been tried, GDPR accepts that organizations may have a
\emph{legitimate interest} in scraping, storing and processing publicly
available PII. In the specific context in which Linkero is designed, we are
talking about an Intellectual Property owner, or an agent in their behalf, that
has a vested interest in gathering information about an online shop that sells
their products, under market value, or without being an officially approved
reseller. Having said that, GDPR definitely limits and restricts, for better or
worse, the ability to extract, store and process public PII.

To conclude this section, it is interesting to mention how ensuring
adequate privacy protection is a difficult balance and has its own trade-offs. This is
the case of the WHOIS protocol maintained by ICANN (i.e. Internet Corporation
for Assigned Names and Numbers), where internet domain registrants' PII are
kept and made public. With the introduction of GDPR, personal details of the
person who registered a domain will have to be hidden. Details of the domain
registrant are valuable information in order to identify the person, or the
signature of the person or group behind specific internet domains. This is
particularly useful when security researchers try to attribute the ownership of
phshing or spamming domains to a particular actor or group. Likewise, in
counterfeit investigations, being able to tell how many and which domains were registered by
the same actor is particularly useful, especially if you are trying to monitor
the actor's activities. Information do not need to be accurate, in fact domain
registrars do not verify the real identity of the person registering a domain.
There are however details that need to be valid, like contact details (e.g.
email address, phone number, \ldots) because on them depends the correct funcioning of
the domain, which is in the registrant's interest. The disappearence of those
PII from WHOIS caused a great deal of frustration in the security community, and
ended up with ICANN filing an official litigation against the European Union
asking for some sort of exhemption in order to keep offering relevant WHOIS
data.

GDPR was created and voted by EU member states in a period when information is
being considered the new oil of the global economy: access to personal data and
consumer behaviour generates huge amounts of revenues for corporations like
Facebook and Google. And the reason for that is simple: targeted ads campaigns
are extremely effective and efficient, and marketing departments are willing to pay
for that. The commercial application of petabyte of consumer data is well known.
But there is another aspect to an unregulated access to data belonging to
billions of people, which is \emph{political influence}. As we have seen during
the 2016 US election, bad actors managed to use personal data of millions of US
citizens to both target specifically voters in swing states, and to analyze
general voters' opinion in order to design very effective slogans. The same
actors fabricated made up stories---literally, fake news---that would appeal to
the most visceral human fears in order to influence undecided voters. That
happended also during the British referendum of Brexit, and was attempted during
the French presidential elections in 2017, and very likely in many other
countries that did not receive as much media attention. At the same time we
saw government bodies abusing their ability to tap into and collect
indiscriminately data, just as Edward Snowden exposed how the US National
Security Agency. GDPR may not be the best solution yet, but it is a much needed
regulation at a time when personal information can have serious impact on the
society.

\section{Server security standards}
A chain is a strong as its weakest link, and a system is as secure as its
most vulnerable element, as the saying goes. In today's reality, any system
facing the open and public internet has to be secured against peole trying to
misuse it. It is not just big banks and multinational corporations that have to
put in place sci-fi grade defenses, but the same private wifi routers and
connected thermostats have to be protected against bad actors that would gladly
take them over and deploy them to perform some fraudulent or criminal
activities. It is well known that professional computer criminals constantly
scan the internet to detect new connected devices and try to add them to their
personal collection of bots for different purposes.
Internet security is therefore a must regardless of the complexity of the
system. Information Security literature identifies three main goals of a good
system security strategy, also referred to as the \emph{CIA triad} \cite{DP00}:
confidentiality, integrity and availability. That means that a system security
strategy aims at guaranteeing that data stored in the system will be accessed
only by legitimate and relevant parties (confidentiality), that the same data
will be kept in its original state away from sources that may alter it,
intercept it or erase it (integrity), and that despite of the safety measures in
place, if an attack succeedes, the data will still be accessible, and business
will continue with minimal to no disruption (availability). In addition to the
CIA triad, some experts also include \emph{accountability}, which essentially
means that if an incident occurs, the remediation team should be able to
identifiy exaclty where the breach happened and the location if not the identity
of the person who executed the attack.

The industry identifies two main types of threat: internal and external.
The external threat is probably the first and most popular to come to mind,
often depicted like the hooded boy wearing a Guy Fawkes mask and typing away on
his laptop: the hacker. Put simply, it is a person or organization that attacks
a system from the outside. The other threat that does not get as much
attention, but can be equally damaging, is the internal threat. The internal
threat is represented by an internal user that intentionally or by mistake
causes damage to the system. An attacker can be further classified based on
his/her level of activity: a \emph{passive} attacker tend to sit and intercept
data exchanges, without affecting the system resources, whereas an \emph{active}
attacker directly alters system resources while the operation is underway.

Any attack normally impacts one or more of system assets: hardware, software,
data or networks.

In conclusion, it is worth mentioning that there are five types of security
measures that an organization can put in place to protect its servers:
deterrent, prevention, detection, response and recovery. Deterrent measures are
meant to scare attackers off before they even consider to start their exploits.
The criminal law with its system of penalties is definitely the first deterrent
that comes to mind. But media articles detailing how sofisticate and complex is
a detection system can offer the same purpose. Preventive measures on the other
end are technologies that anticipate potential attacks and stop them before any
damage occurs. Firewalls are a great example. Since absolute prevention
is rarely feasible, detection technologies can cover those situation in
which an attack is found while still ongoing. Detection immediately
triggers response measures, such as blacklisting an IP associated with a
number of failed login attempts. Finally, recovery measures are the ones that
are deployed after an attack occurred, and try to assess the damage and
remediate where possible.

Information security is a very vast and complex discipline, which
directly reflects the complexity of modern interconnected systems. For the
purpose of this project however we will only consider the basic security
configuration required to protect a standalone linux server.

In the next part of this section we will review in greater details some of the
prevention technologies and solutions available, particularly those that were
considered for this project, focussing on prevention, detection, response and
recovery measures.

\subsection{Prevention}

\subsubsection{Regular software updates}
Asset: software
Attack type: active
Attacker: external
Applications running on the system are rarely bug-free. It is always good
practice to keep every system application up-to-date with latest patches, since
every time a new vulnerability is discovered, application developers tend
to issue a new version or patch to close it.

\subsubsection{Environment isolation}
Rather than a technology, \emph{isolation} is a principle and a policy that
recommends every resource should be isolated from the rest, and access between
them should be regulated and monitored. This is particularly critical in the
context of systems that have public interfaces. Isolation applies to file
management as well, requiring users files and folders to be isolated from one
another, preventing one user's compromised credentials to be used to access
everyone else's files. Finally, isolation applies to the same security
mechanisms, since they should be isolated from the rest of the system, so that
they cannot be compromised by the attacker.

\subsubsection{Virtual Private Netwroks}
VPN are a good solution to connect parts of the system that are not located on
the same machine, or in the same network. Since communication between resources
has to go through the public network, VPN offer a way to encrypt and protect
data exchange based on asymmetric encryption technology.

\subsubsection{SSL and TLS encryption}
Similarly to VPN, SSL and TLS are two technologies that create an encrypted
communication channel between a client and a server. The most common use case
for these solutions is when login forms are required, so that a client can
transfer login credentials to the server using a HTTPS connection, thus avoiding
to transfer username and password in clear over a non-secure public network.

\subsubsection{SSH}
The Secure Shell, SSH, is the default application for remote server management.
Like VPN and SSL, it establishes an encrypted connection between a client and a
server, relying on asymmetric encryption. It allows system administrators to
login remotely to the server command line shell. It can be configured to either
accept a standard password login, or to accept autmoatic logins from approved
clients only.

\subsubsection{Firewalls}
Firewalls are also considered \emph{intrusion prevention systems}.

\subsubsection{Users and groups least privileges}

\subsubsection{Password policy}

\subsubsection{Encryption of data at rest}

\subsubsection{Physical security}

\subsection{Detection}

\subsection{Response}

\subsection{Recovery}



\subsection{Intrusion detection systems}

\subsection{Regular file, service and vulnerability audits}

\subsection{Business continuity plan}
Backups, duplication\ldots



\section{Ergonomy of software interface}
Loren Ipsum \ldots

\section{Code development principles}
Robert Martin S.O.L.I.D. development principles.

\section{Existing brand monitoring services}
Loren Ipsum \ldots